Code used to generate paper results : Database comparison and CT characteristics

### 2024-05-17 Paper_code-CT_analysis.Rmd
Vefified and corrected all figs

```{r library, message = FALSE}
library(readxl)
library(readr)
library(CTexploreR)
library(tidyverse)
library(SummarizedExperiment)
library(ComplexHeatmap)
library(circlize)
library(SingleCellExperiment)
library(msigdbr)
library(DOSE)
library(org.Hs.eg.db)
library(clusterProfiler)
library(patchwork)
library(BiocParallel)
library(Biostrings)
library(AUCell)
```

@@

@@

Chr_colors = c("X-linked" = "deeppink", "not X" = "lightgreen")

@@

Fetal_cell_colors = c( "F_PGC" = "pink", 
                       "F_GC" = "pink4", 
                       "F_oogonia" = "palevioletred3",
                       "F_oocyte" = "mediumorchid4", 
                       "M_PGC" = "lightblue1", 
                       "M_GC" = "steelblue3",
                       "M_pre_spermatogonia" = "darkblue")

oocytes_colors <- c("Growing oocytes" = "paleturquoise3", 
                  # "Fully grown oocytes" ="lightgreen", 
                  # "Metaphase I" = "mediumvioletred", 
                  "Metaphase II" = "mediumvioletred")

sex_colors <- c("F" = "palevioletred1", "M"= "skyblue2")

stage_colors <- c( "pre-meiotic" = "moccasin", 
                   "meiotic"= "lightsalmon",
                   "post-meiotic" = "maroon4",
                   "somatic" = "gray")

@@

Function to print GTEX expression with specific annotations.
```{r GTEX_function}
print_GTEX_heatmap <- function(genes = NULL, units = "log_TPM", 
                               row_fontsize = 2,
                               column_fontsize = 6, 
                               h_width = 3,
                               h_height = 10,
                               split_rows_by = "regulation") {

@@

```{r meth_value_heatmap}
met_genes <- CT_genes %>% 
  filter(regulated_by_methylation) %>% 
  pull(external_gene_name)
met_value_number_per_gene <- tibble(
  gene = CT_genes$external_gene_name,
  regulated_by_methylation = CT_genes$regulated_by_methylation,
  NA_met_values_number = rowSums(
    is.na(assay(mean_methylation_in_tissues[CT_genes$external_gene_name, ]))))
n_met_values <- ncol(mean_methylation_in_tissues)
CT_met_with_WGBS <- met_value_number_per_gene %>%
  filter(regulated_by_methylation) %>% 
  filter(NA_met_values_number !=  n_met_values)

@@

```{r DAC_heatmap}
DAC_fig <- function(genes = gene_type$gene, multimapping = TRUE, 
                    row_fontsize = 2,
                    units = "log1p",
                    values_only = FALSE, scale_lims = NULL) {

@@

  column_ha_treatment <- HeatmapAnnotation(
    treatment = df_col$treatment,
    col = list(treatment = c("CTL" = "royalblue4", "5-aza" = "maroon3")),
    annotation_legend_param = legends_param,
    border = FALSE, 
    simple_anno_size = unit(0.2, "cm"),
    annotation_name_gp = gpar(fontsize = 0))

@@

h_DAC <- DAC_fig(CT_genes$external_gene_name, 
                 multimapping = TRUE, 
                 units = "log1p",
                 row_fontsize = fs
)

@@

```{r prop_meth}
x <- enframe(table(CT_genes_types$Regulation), name = "Regulation") %>% 
  left_join(enframe(prop.table(table(CT_genes_types$Regulation)), 
                    name = "Regulation", value = "Percent"))
met_prop_fig <- ggplot(x, aes(x = '', y = value, fill = Regulation)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start = 0) +
  geom_text(aes(y = value/5 + c(0, cumsum(value)[-length(value)]), 
                label = rev(paste0(Regulation, "\n", value, "/", nrow(CT_genes_types), "\n",
                               round(Percent * 100), "%"))))+
            #, size = 6
  scale_fill_manual(values = c( "indianred2", "cyan4")) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none")
met_prop_fig
```

@@

fig_X_enrichment <- ggplot(CT_genes_types, aes(x = chr, fill = Regulation)) +
  geom_bar(stat = 'count') +
  xlab("Chromosome") +
  ylab("Gene number") +
  scale_fill_manual(values = Regulation_colors) +
  theme(legend.position = "bottom",
        aspect.ratio = 1)
fig_X_enrichment
```

@@

```{r meth_figs}
x <- enframe(table(CT_genes_types$Regulation), name = "Regulation") %>% 
  left_join(enframe(prop.table(table(CT_genes_types$Regulation)), 
                    name = "Regulation", value = "Percent"))
prop_fig <- ggplot(x, aes(x = '', y = value, fill = Regulation)) +
  geom_bar(stat="identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(y = value/5 + c(0, cumsum(value)[-length(value)]),
                label = rev(paste0(Regulation, "\n", value, "/", nrow(CT_genes_types), "\n",
                               round(Percent * 100), "%"))), size = 1.2 )+
  scale_fill_manual(values = Regulation_colors) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none")

fig_freq_by_chr <- ggplot(CT_genes_types, aes(x = chr, fill = Regulation)) +
  geom_bar(stat = 'count') +
  xlab("Chromosome") +
  ylab("Gene number") +
  scale_fill_manual(values = Regulation_colors) +
  theme_test()+
  theme(legend.position = "top",
        aspect.ratio = 1,
        axis.text = element_text(size = 3),
        axis.title = element_text(size = 5),
        axis.ticks = element_line(linewidth = 0.2),
        axis.ticks.length = unit(0.5, "mm"),
        legend.text = element_text(size = 5),
        legend.key.size = unit(2, "mm"),
        legend.title = element_blank())

@@

# combine with CCLE data
load("~/cluster/Packages/CTdata/eh_data/CCLE_data.rda")
ccle <- CCLE_data
tumors <- unique(ccle$primary_disease)

@@

fig_mean_prop_in_tum <- ggplot(mean_and_prop,
                               aes(x = mean_log, y = proportion_of_positive_tum,
                                   shape = X_linked,
                                   color = Regulation,
                                   label = gene)) +
  geom_point(size = 1) +
  scale_colour_manual(values = Regulation_colors)+
  ggrepel::geom_text_repel(
    data = mean_and_prop %>% 
      filter(proportion_of_positive_tum > 20 | 
               (proportion_of_positive_tum > 5 & mean_log > 4)), 
    aes(x = mean_log, y = proportion_of_positive_tum, 
        label = gene), size = 1.5, color = "black", min.segment.length = 1,
    max.overlaps = 30, direction = "both") +
  xlab("Mean expression log(TPM + 1)\nin positive samples") +
  ylab("Proportion of positive samples (%)") +
  facet_wrap(~ source, ncol = 1) +
  theme_bw() +
  theme(legend.position = "right",
        aspect.ratio = 0.8,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.3, 'cm'),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 7),
        axis.text = element_text(size = 5),
        axis.ticks = element_line(size = 0.1),
        axis.title = element_text(size = 6),
        strip.text = element_text(size = 6))

@@

  h <- Heatmap(mat[gene_type$gene, rownames(df_col)],
               name = "log(TPM +1)",
               column_title = levels(df_col$project_id),
               column_title_gp = gpar(col =c("COAD" = "darkorchid2", 
                                             "HNSC" = "deeppink2",
                                             "LUSC" = "seagreen3", 
                                             "SKCM" = "red3"), 
                                      fontsize = 3, face = "bold"),
               clustering_method_rows = clust_method,
               clustering_method_columns = clust_method,
               column_split = split_by,
               cluster_column_slices = FALSE,
               row_gap = unit(c(1), "mm"),
               column_gap = unit(c(1), "mm"),
               col = colorRamp2(seq(scale_lims[1], scale_lims[2], length = 11),
                                legend_colors),
               cluster_rows = clustRow,
               cluster_columns = clustCol,
               show_row_names = TRUE,
               show_column_names = FALSE,
               show_row_dend = FALSE,
               show_column_dend = FALSE,
               column_names_gp = gpar(fontsize = 4),
               column_names_centered = TRUE,
               row_names_gp = gpar(fontsize = fontsize),
               row_names_side = "left",
               border = TRUE,
               border_gp = gpar(lwd = 0.05),
               column_names_side = c("top"),
               column_names_rot = 90,
               row_title_gp = gpar(fontsize = 0),
               top_annotation = topAnnot,
               row_split = rowSplit,
               cluster_row_slices = FALSE,
               heatmap_width = unit(h_width, "cm"),
               heatmap_height = unit(h_height, "cm"),
               use_raster = TRUE,
               raster_device = "CairoPNG",
               raster_quality = 10,
               heatmap_legend_param = legends_param_TCGA,
               show_heatmap_legend = TRUE)

@@

h1 <- print_TCGA_heatmap_with_AUCscores(genes = met_genes, 
                                        clustCol = TRUE,
                                        clust_method = "ward.D",
                                        h_width = 6,
                                        h_height = 5,
                                        fontsize = 1, 
                                        scale_lims = c(0, 6),
                                        # tum_type = c("SKCM", "LUAD", "COAD", "BRCA"),
                                        tum_type = tum_types,
                                        split_rows_by = NULL)
draw(h1, column_title = "CT genes regulated by methylation",
     column_title_gp = gpar(fontsize = 4), 
     merge_legend = TRUE)

@@

h2 <- print_TCGA_heatmap_with_AUCscores(genes = not_met_genes, 
                                       clustCol = TRUE,
                                       AUScores = "AUC_genes_not_met", 
                                       show_AUScore = TRUE,
                                       clust_method = "ward.D",
                                       h_width = 6,
                                       h_height = 2.5,
                                       fontsize = 1, 
                                       scale_lims = c(0, 6),
                                       #  tum_type = c("SKCM", "LUAD", "COAD", "BRCA"),
                                       tum_type = tum_types,
                                       split_rows_by = NULL)
draw(h2, column_title = "CT genes not regulated by methylation",
     column_title_gp = gpar(fontsize = 4))

@@

f1 <- as_tibble(colData(TCGA), rownames = "barcode") %>% 
  filter(project_id %in% tum_types) %>% 
  mutate(project_id = factor(project_id, levels = tum_types,)) %>% 
  ggplot(aes(x = global_methylation, y = AUC_genes_regulated_by_met, color = project_id)) +
  geom_point(size = 0.3) +
  facet_wrap(~ project_id, nrow = 1) +
  scale_color_manual(values = c("COAD" = "darkorchid2", "HNSC" = "deeppink2",
                                "LUSC" = "seagreen3", "SKCM" = "red3")) + 
  geom_smooth(method = 'lm', color = "black", linewidth = 0.3, se = FALSE) +
  xlab('Global methylation level') +
  ylab("AUCscore CT genes\nregulated by methylation") +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 9),
        axis.text = element_text(size = 4),
        axis.ticks = element_line(size = 0.1)) +
  ggpubr::stat_cor(method="pearson", size = 2.5, label.x = .4, 
           label.y = 0.11, label.sep = "\n", color = "black") +
  ylim(0, 0.12)

f2 <- as_tibble(colData(TCGA), rownames = "barcode") %>% 
  filter(project_id %in% tum_types) %>% 
  mutate(project_id = factor(project_id, levels = tum_types,)) %>% 
  ggplot(aes(x = global_methylation, y = AUC_genes_not_regulated_by_met, color = project_id)) +
  geom_point(size = 0.3) +
  facet_wrap(~ project_id, nrow = 1) +
  scale_color_manual(values = c("COAD" = "darkorchid2", "HNSC" = "deeppink2",
                                "LUSC" = "seagreen3", "SKCM" = "red3")) + 
  geom_smooth(method = 'lm', color = "black", linewidth = 0.3, se = FALSE) +
  xlab('Global methylation level') +
  ylab("AUCscore CT genes\nnot regulated by methylation") +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 9),
        axis.text = element_text(size = 4),
        axis.ticks = element_line(size = 0.1)) +
  ggpubr::stat_cor(method="pearson", size = 2.5, label.x = .4, 
           label.y = 0.11, label.sep = "\n", color = "black") +
  ylim(0, 0.12)

@@

pdf(file = paste0("figs/CT_not_met_AUCscore_hypo_TCGA.pdf"), height = 2.5)
f2
dev.off()
```
# CT expression in healthy cells (fetal and adult)

@@

```{r}
################################################################################
# Load testis scRNAseq data
################################################################################

@@

#######################################################################
## Function to draw expression in adult testis cells with specific annotations
## and downasampling the number of cells from each cell type
#######################################################################
draw_testis_heatmap <- function(genes, 
                                split_by = "chr_met", 
                                sex = c("F", "M"),
                                ncells_max = NULL,
                                font_size = 6,
                                scale_lims = NULL,
                                clusterRow = FALSE,
                                show_legend = TRUE,
                                h_width = 5,
                                h_height = 10,
                                clust_method = "ward.D"){

@@

######################################################################
## Function to draw expression in adult oocytes cells with specific annotations
## and downasampling the number of cells from each cell type
#######################################################################
draw_oocytes_heatmap <- function(genes, 
                               split_by = "chr_met", 
                               ncells_max = NULL,
                               font_size = 6,
                               scale_lims = NULL,
                               clusterRow = TRUE,
                               show_legend = TRUE,
                               h_width = 5,
                               h_height = 10,
                               clust_method = "ward.D"){

@@

  column_ha_cell_type = HeatmapAnnotation(
    oocytes_cell_type = df_col_oocytes$type,
    col = list(Adult_cell_type = oocytes_colors),      
    border = TRUE,
    simple_anno_size = unit(0.15, "cm"),
    annotation_name_gp = gpar(fontsize = 0),
    annotation_legend_param = legends_param_sc)

>>>>

```{r library, message = FALSE}
library(readxl)
library(readr)
library(CTexploreR)
library(tidyverse)
library(SummarizedExperiment)
library(ComplexHeatmap)
library(circlize)
library(SingleCellExperiment)
library(msigdbr)
library(DOSE)
library(org.Hs.eg.db)
library(clusterProfiler)
library(patchwork)
library(AUCell)
```

@@

DAC_treatment_colors <- c("CTL" = "royalblue4", "5-aza" = "maroon3")

@@

Chr_colors = c("X-linked" = "deeppink3", "not X" = "lightgreen")

@@

Fetal_cell_colors = c( "F_PGC" = "pink", 
                       "F_GC" = "pink3", 
                       "F_oogonia" = "palevioletred3",
                       "F_oocyte" = "mediumorchid4", 
                       "M_PGC" = "lightblue1", 
                       "M_GC" = "steelblue3",
                       "M_pre_spermatogonia" = "royalblue3")

oocytes_colors <- c("Growing oocytes" = "mediumpurple1",
                    # "Fully grown oocytes" ="lightgreen",
                     # "Metaphase I" = "mediumvioletred",
                    "Metaphase II" = "mediumpurple4")

sex_colors <- c("F" = "palevioletred1", "M"= "skyblue1")

stage_colors <- c( "pre-meiotic" = "peachpuff", 
                   "meiotic"= "lightsalmon",
                   "post-meiotic" = "maroon4",
                   "somatic" = "gray")

@@

Function to print GTEX expression with specific annotations.

```{r GTEX_function}
print_GTEX_heatmap <- function(genes = NULL, units = "log_TPM", 
                               row_fontsize = 2,
                               column_fontsize = 6, 
                               h_width = 3,
                               h_height = 10,
                               split_rows_by = "regulation") {

@@

```{r meth_heatmap_function}
met_genes <- CT_genes %>% 
  filter(regulated_by_methylation) %>% 
  pull(external_gene_name)
met_value_number_per_gene <- tibble(
  gene = CT_genes$external_gene_name,
  regulated_by_methylation = CT_genes$regulated_by_methylation,
  NA_met_values_number = rowSums(
    is.na(assay(mean_methylation_in_tissues[CT_genes$external_gene_name, ]))))
n_met_values <- ncol(mean_methylation_in_tissues)
CT_met_with_WGBS <- met_value_number_per_gene %>%
  filter(regulated_by_methylation) %>% 
  filter(NA_met_values_number !=  n_met_values)

@@

```{r DAC_heatmap_function}
DAC_fig <- function(genes = gene_type$gene, multimapping = TRUE, 
                    row_fontsize = 2,
                    units = "log1p",
                    values_only = FALSE, scale_lims = NULL) {

@@

  column_ha_treatment <- HeatmapAnnotation(
    treatment = df_col$treatment,
    col = list(treatment = DAC_treatment_colors),
    annotation_legend_param = legends_param,
    border = FALSE, 
    simple_anno_size = unit(0.2, "cm"),
    annotation_name_gp = gpar(fontsize = 0))

@@

h_DAC <- DAC_fig(CT_genes$external_gene_name, 
                 multimapping = TRUE, 
                 units = "log1p",
                 row_fontsize = fs)

@@

```{r prop_meth}
x <- enframe(table(CT_genes_types$Regulation), name = "Regulation") %>% 
  left_join(enframe(prop.table(table(CT_genes_types$Regulation)), 
                    name = "Regulation", value = "Percent"))
met_prop_fig <- ggplot(x, aes(x = '', y = value, fill = Regulation)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start = 0) +
  geom_text(aes(y = value/5 + c(0, cumsum(value)[-length(value)]), 
                label = rev(paste0(Regulation, "\n", value, "/", 
                                   sum(table(CT_genes_types$Regulation)), "\n",
                               round(Percent * 100), "%"))))+
            #, size = 6
  scale_fill_manual(values = c( "indianred2", "cyan4")) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none")
met_prop_fig
```

@@

fig_X_enrichment <- ggplot(CT_genes_types[
  !is.na(CT_genes_types$regulated_by_methylation), ], 
                           aes(x = chr, fill = Regulation)) +
  geom_bar(stat = 'count') +
  xlab("Chromosome") +
  ylab("Gene number") +
  scale_fill_manual(values = Regulation_colors) +
  theme(legend.position = "bottom",
        aspect.ratio = 1)
fig_X_enrichment
```

@@

```{r meth_figs}
x <- enframe(table(CT_genes_types$Regulation), name = "Regulation") %>% 
  left_join(enframe(prop.table(table(CT_genes_types$Regulation)), 
                    name = "Regulation", value = "Percent"))
prop_fig <- ggplot(x, aes(x = '', y = value, fill = Regulation)) +
  geom_bar(stat="identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(y = value/5 + c(0, cumsum(value)[-length(value)]),
                label = rev(paste0(Regulation, "\n", value, "/", 
                                   sum(table(CT_genes_types$Regulation)), "\n",
                               round(Percent * 100), "%"))), size = 1.2 )+
  scale_fill_manual(values = Regulation_colors) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "none")

fig_freq_by_chr <- ggplot(CT_genes_types[
  !is.na(CT_genes_types$regulated_by_methylation), ], 
  aes(x = chr, fill = Regulation)) +
  geom_bar(stat = 'count') +
  xlab("Chromosome") +
  ylab("Gene number") +
  scale_fill_manual(values = Regulation_colors) +
  theme_test()+
  theme(legend.position = "top",
        aspect.ratio = 1,
        axis.text = element_text(size = 3),
        axis.title = element_text(size = 5),
        axis.ticks = element_line(linewidth = 0.2),
        axis.ticks.length = unit(0.5, "mm"),
        legend.text = element_text(size = 5),
        legend.key.size = unit(2, "mm"),
        legend.title = element_blank())

@@

# combine with CCLE data
ccle <- CCLE_data
tumors <- unique(ccle$primary_disease)

@@

fig_mean_prop_in_tum <- ggplot(mean_and_prop[!is.na(mean_and_prop$Regulation), ],
                               aes(x = mean_log, y = proportion_of_positive_tum,
                                   shape = X_linked,
                                   color = Regulation,
                                   label = gene)) +
  geom_point(size = 1) +
  scale_colour_manual(values = Regulation_colors)+
  ggrepel::geom_text_repel(
    data = mean_and_prop %>% 
      filter(proportion_of_positive_tum > 20 | 
               (proportion_of_positive_tum > 5 & mean_log > 4)), 
    aes(x = mean_log, y = proportion_of_positive_tum, 
        label = gene), size = 1.5, color = "black", min.segment.length = 1,
    max.overlaps = 30, direction = "both") +
  xlab("Mean expression log(TPM + 1)\nin positive samples") +
  ylab("Proportion of positive samples (%)") +
  facet_wrap(~ source, ncol = 1) +
  theme_bw() +
  theme(legend.position = "right",
        aspect.ratio = 0.8,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.3, 'cm'),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 7),
        axis.text = element_text(size = 5),
        axis.ticks = element_line(size = 0.1),
        axis.title = element_text(size = 6),
        strip.text = element_text(size = 6))

@@

  h <- Heatmap(mat[gene_type$gene, rownames(df_col)],
               name = "log(TPM +1)",
               column_title = levels(df_col$project_id),
               column_title_gp = gpar(col = TCGA_colors[tum_type], 
                                      fontsize = 3, face = "bold"),
               clustering_method_rows = clust_method,
               clustering_method_columns = clust_method,
               column_split = split_by,
               cluster_column_slices = FALSE,
               row_gap = unit(c(1), "mm"),
               column_gap = unit(c(1), "mm"),
               col = colorRamp2(seq(scale_lims[1], scale_lims[2], length = 11),
                                legend_colors),
               cluster_rows = clustRow,
               cluster_columns = clustCol,
               show_row_names = TRUE,
               show_column_names = FALSE,
               show_row_dend = FALSE,
               show_column_dend = FALSE,
               column_names_gp = gpar(fontsize = 4),
               column_names_centered = TRUE,
               row_names_gp = gpar(fontsize = fontsize),
               row_names_side = "left",
               border = TRUE,
               border_gp = gpar(lwd = 0.05),
               column_names_side = c("top"),
               column_names_rot = 90,
               row_title_gp = gpar(fontsize = 0),
               top_annotation = topAnnot,
               row_split = rowSplit,
               cluster_row_slices = FALSE,
               heatmap_width = unit(h_width, "cm"),
               heatmap_height = unit(h_height, "cm"),
               use_raster = TRUE,
               raster_device = "CairoPNG",
               raster_quality = 10,
               heatmap_legend_param = legends_param_TCGA,
               show_heatmap_legend = TRUE)

@@

h1 <- print_TCGA_heatmap_with_AUCscores(genes = met_genes, 
                                        clustCol = TRUE,
                                        clust_method = "ward.D",
                                        h_width = 6,
                                        h_height = 5,
                                        fontsize = 1, 
                                        scale_lims = c(0, 6),
                                        tum_type = tum_types,
                                        split_rows_by = NULL)
draw(h1, column_title = "CT genes regulated by methylation",
     column_title_gp = gpar(fontsize = 4), 
     merge_legend = TRUE)

@@

h2 <- print_TCGA_heatmap_with_AUCscores(genes = not_met_genes, 
                                       clustCol = TRUE,
                                       AUScores = "AUC_genes_not_met", 
                                       show_AUScore = TRUE,
                                       clust_method = "ward.D",
                                       h_width = 6,
                                       h_height = 2.5,
                                       fontsize = 1, 
                                       scale_lims = c(0, 6),
                                       tum_type = tum_types,
                                       split_rows_by = NULL)
draw(h2, column_title = "CT genes not regulated by methylation",
     column_title_gp = gpar(fontsize = 4))

@@

f1 <- as_tibble(colData(TCGA), rownames = "barcode") %>% 
  filter(project_id %in% tum_types) %>% 
  mutate(project_id = factor(project_id, levels = tum_types,)) %>% 
  ggplot(aes(x = global_methylation, y = AUC_genes_regulated_by_met, 
             color = project_id)) +
  geom_point(size = 0.3) +
  facet_wrap(~ project_id, nrow = 1) +
  scale_color_manual(values = TCGA_colors) + 
  geom_smooth(method = 'lm', color = "black", linewidth = 0.3, se = FALSE) +
  xlab('Global methylation level') +
  ylab("AUCscore CT genes\nregulated by methylation") +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 9),
        axis.text = element_text(size = 4),
        axis.ticks = element_line(size = 0.1)) +
  ggpubr::stat_cor(method="pearson", size = 2.5, label.x = .4, 
           label.y = 0.11, label.sep = "\n", color = "black") +
  ylim(0, 0.12)

f2 <- as_tibble(colData(TCGA), rownames = "barcode") %>% 
  filter(project_id %in% tum_types) %>% 
  mutate(project_id = factor(project_id, levels = tum_types,)) %>% 
  ggplot(aes(x = global_methylation, y = AUC_genes_not_regulated_by_met, 
             color = project_id)) +
  geom_point(size = 0.3) +
  facet_wrap(~ project_id, nrow = 1) +
  scale_color_manual(values = TCGA_colors) + 
  geom_smooth(method = 'lm', color = "black", linewidth = 0.3, se = FALSE) +
  xlab('Global methylation level') +
  ylab("AUCscore CT genes\nnot regulated by methylation") +
  theme(legend.position = "none",
        aspect.ratio = 1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 9),
        axis.text = element_text(size = 4),
        axis.ticks = element_line(size = 0.1)) +
  ggpubr::stat_cor(method="pearson", size = 2.5, label.x = .4, 
           label.y = 0.11, label.sep = "\n", color = "black") +
  ylim(0, 0.12)

@@

pdf(file = paste0("figs/CT_not_met_AUCscore_hypo_TCGA.pdf"), height = 2.5)
f2
dev.off()
```

# CT expression in healthy cells (fetal and adult)

@@

```{r, testis_data}
################################################################################
# Load testis scRNAseq data
################################################################################

@@

################################################################################
## Function to draw expression in adult testis cells with specific annotations
## and downasampling the number of cells from each cell type
################################################################################
draw_testis_heatmap <- function(genes, 
                                split_by = "chr_met", 
                                sex = c("F", "M"),
                                ncells_max = NULL,
                                font_size = 6,
                                scale_lims = NULL,
                                clusterRow = FALSE,
                                show_legend = TRUE,
                                h_width = 5,
                                h_height = 10,
                                clust_method = "ward.D"){

@@

################################################################################
## Function to draw expression in adult oocytes cells with specific annotations
## and downasampling the number of cells from each cell type
################################################################################
draw_oocytes_heatmap <- function(genes, 
                               split_by = "chr_met", 
                               ncells_max = NULL,
                               font_size = 6,
                               scale_lims = NULL,
                               clusterRow = TRUE,
                               show_legend = TRUE,
                               h_width = 5,
                               h_height = 10,
                               clust_method = "ward.D"){

@@

  column_ha_cell_type = HeatmapAnnotation(
    Oocytes_cell_type = df_col_oocytes$type,
    col = list(Oocytes_cell_type = oocytes_colors),      
    border = TRUE,
    simple_anno_size = unit(0.15, "cm"),
    annotation_name_gp = gpar(fontsize = 0),
    annotation_legend_param = legends_param_sc)
